\documentclass[11pt,a4paper,sans]{moderncv}

\moderncvstyle{banking}
\moderncvcolor{black}
\nopagenumbers{}

\usepackage[utf8]{inputenc}
\usepackage[
    top=1.5cm,
    bottom=1.5cm,
    left=1.5cm,
    right=1.5cm
]{geometry}
\usepackage{enumitem}

\name{Rolando Palermo}{Rodríguez Cruz}

\begin{document}

\makecvtitle

\vspace*{-10mm}

\begin{center}
\textbf{Lead Software Engineer | Backend \& Cloud Architect}
\end{center}

\begin{center}
\begin{tabular}{c c c}
+52 4451401994 & rolando.roc@gmail.com & Guanajuato, México \\
linkedin.com/in/RolandoPalermo & github.com/RolandoPalermo &
\end{tabular}
\end{center}

\section{Profile}
Lead Software Engineer with 13+ years of experience building backend systems and cloud-based applications. Strong focus on Java, Kotlin, and Spring Boot. Experienced working on real-world systems in retail, logistics, payments, and electronic invoicing.

\section{Professional Experience}

\cventry{Aug 2025 -- Present}{Lead Software Engineer}{MHP -- A Porsche Company}{Guadalajara, Mexico}{}{
\begin{itemize}
\item Leading the design of a microservices architecture for a VDA5050/LIF-based logistics platform, helping the system scale better and making it easier to maintain.
\item Led a cross-functional engineering team of six developers (3 backend, 2 frontend) in Mexico while coordinating architecture, delivery, and technical alignment with three distributed teams across Germany (2) and Romania (1).
\item Own technical backlog decisions and collaborate with Product Management on roadmap and architecture.
\item Drive engineering best practices through code reviews, mentoring, and sprint leadership.
\item Design REST APIs, persistence layers, and messaging integrations using Kotlin 2.2, Java 21, Spring Boot 3, Jetbrains Exposed, PostgreSQL, MQTT, and RabbitMQ.
\item Developed interactive web applications using React, TypeScript, and React Flow, building reusable functional components to create graph-based editors for logistics maps and robot navigation workflows.
\item Define and implement authentication and authorization strategies based on OAuth 2.0, OpenID Connect (OIDC), JWT, Client Credentials, and Authorization Code Flow with PKCE for secure machine-to-machine and user-facing integrations.
\item Designed and implemented an AI-powered support assistant using LLMs and Retrieval-Augmented Generation (RAG), enabling users to ask natural-language questions about robot status, diagnostics, and technical documentation, significantly reducing the need to create support tickets.
\item Designed and automated cloud infrastructure provisioning and server configuration workflows using Terraform and Ansible.
\end{itemize}}

\cventry{Sep 2024 -- Jul 2025}{Technical Lead (Promoted from Lead Software Engineer)}{Petco}{Queretaro, Mexico}{}{
\begin{itemize}
\item Promoted to Technical Lead to work on hardware and software initiatives across retail operations in 1,000+ stores, helping improve system reliability and day-to-day store operations.
\item Managed and mentored three cross-functional engineering teams (two in Mexico and one in India), leading a total of eight engineers while coordinating technical delivery across distributed teams.
\item Led integrations with payment industry platforms (Ingenico, Verifone, ACI), ensuring secure, reliable, and high-volume transaction processing across retail locations.
\item Designed and implemented REST APIs following microservices architecture principles for a Point of Sale (POS) platform using Java 11, Java 17, Spring Boot 3, enabling scalable and distributed system integrations.
\item Built cloud-native solutions on AWS (S3, EKS), improving scalability and deployment flexibility.
\item Optimized data workflows and system integrations across PostgreSQL and SQL Server, improving performance and data consistency.
\item Worked within Agile/Scrum teams, delivering backend features through iterative development and continuous improvement practices.
\end{itemize}}

\cventry{Jul 2024 -- Sep 2024}{Shopfloor Software Specialist}{Forvia}{Tlaquepaque, Mexico}{}{
\begin{itemize}
\item Developed and maintained manufacturing and shopfloor applications using Angular 13, TypeScript, ng-zorro, Java 11, Java 17, Spring Boot 3, and MySQL, supporting production operations and business workflows.
\item Built responsive and user-friendly interfaces using Angular 13, TypeScript, ng-zorro, HTML, and CSS, improving usability and reducing user interaction friction.
\item Developed reusable Angular components, TypeScript services, and reactive forms to streamline data entry, validation, and workflow management across multiple modules.
\item Implemented API Integration between Angular frontend applications and Spring Boot microservices using REST APIs, enabling real-time communication between shopfloor systems and backend services.
\item Consumed and integrated REST APIs using Angular HttpClient and TypeScript, handling authentication, error management, request/response mapping, and business data processing.
\item Partnered with cross-functional Agile/Scrum teams to deliver backend solutions in iterative development cycles.
\item Automated build, testing, and deployment pipelines using GitLab CI/CD, increasing release reliability and accelerating delivery cycles.
\end{itemize}}

\cventry{Jul 2021 -- May 2024}{Backend Developer}{Rappi}{Mexico City, Mexico}{}{
\begin{itemize}
    \item Worked on an event-driven microservices architecture using Java 8, Java 11, Kotlin 1.6, Spring Boot 3, REST APIs, PostgreSQL and Kafka, helping the system handle higher traffic and making integrations more reliable.
\item Applied multithreading, concurrency, and performance optimization techniques to design and build scalable, high-performance applications, improving responsiveness and resource efficiency.
    \item Built and maintained data pipelines using DBT, Airflow, Astronomer and Snowflake supporting reliable data processing and analytics workflows.
    \item Implemented and optimized CI/CD pipelines with Docker, Jenkins, and Groovy, increasing deployment automation and release consistency.
    \item Ensured code quality through unit testing with JUnit and Mockito, and continuous inspection with SonarQube.
\item Participated in Agile/Scrum ceremonies, helping deliver reliable backend services through iterative development and continuous refinement.
\end{itemize}}

\cventry{Jul 2019 -- Apr 2021}{Full Stack Engineer}{BairesDev}{Buenos Aires, Argentina}{}{
\begin{itemize}
\item Promoted from Backend Engineer to Full Stack Engineer to lead end-to-end feature development across backend services and frontend applications.
\item Designed and implemented scalable REST APIs using Java 8, Java 11, Spring Boot 2.7, Spring Security, OAuth 2.0, JPA/Hibernate, MySQL, Flyway, and Gradle, enabling secure and reliable system integrations.
\item Designed and implemented responsive user interfaces using Angular, Angular Material, TypeScript, HTML, CSS, and RxJS, delivering rich client-side functionality and improved application performance.
\item Implemented logging and monitoring solutions using Log4j and Splunk, enhancing system observability and reducing issue resolution time.
\item Built and maintained CI/CD pipelines using Jenkins, increasing deployment automation and release consistency.
\item Ensured code quality through comprehensive unit testing using JUnit (backend) and Jasmine/Karma (frontend).
\item Collaborated in Agile/Scrum teams using Jira and Confluence to deliver features iteratively and efficiently.
\end{itemize}}

\cventry{May 2018 -- May 2019}{Java Backend Developer}{Softtek}{California, USA}{}{
\begin{itemize}
    \item Developed backend services and enterprise integrations using Java 8, Spring Boot 2, and BPMN 2.0 with Camunda, supporting complex business workflows.
    \item Built and maintained SOAP web services and REST APIs, working with XML technologies such as XSLT, XPath, XQuery, and JAXB for data transformation and integration.
    \item Designed and optimized data processing and persistence using Oracle 11g, MySQL, and Apache Ignite.
    \item Worked on document generation solutions using Perceptive Document Composition, supporting business reporting and output workflows.
    \item Collaborated in Agile environments (Scrum/Kanban) and used tools such as Jira, Bamboo, Jenkins, GitLab, and Confluence to support development and delivery processes.
\end{itemize}}

\cventry{Nov 2017 -- Apr 2018}{Java Senior Developer}{Seguritech}{León, Mexico}{}{
\begin{itemize}
    \item Designed and developed computer-aided dispatch (CAD) systems to support real-time operations and improve coordination efficiency.
    \item Built dynamic and responsive user interfaces using AngularJS, jQuery, JavaScript, HTML5, and CSS.
    \item Developed and integrated backend services using Java 8, Apache CXF, Spring (Core, MVC, REST, JPA), and JAX-WS, enabling reliable system communication.
    \item Optimized complex SQL queries, improving database performance and reducing response times.
\item Contributed to the successful delivery of backend services by collaborating in Agile/Scrum environments and embracing iterative development.
\end{itemize}}

\cventry{Mar 2017 -- Aug 2017}{Java Developer}{MDP}{Lima, Peru}{}{
\begin{itemize}
    \item Led the migration of a legacy application from Struts 2 to Spring MVC 4 using Java 8 and SQL Server.
    \item Replaced a boilerplate authentication model with a Spring Security-based solution, implementing a flexible authentication framework supporting multiple identity providers.
    \item Worked within Agile environments (Scrum and Kanban), contributing to iterative delivery and continuous process improvement.
\end{itemize}}

\cventry{Nov 2016 -- Feb 2017}{Java Developer}{IA Interactive}{Morelia, Mexico}{}{
\begin{itemize}
    \item Developed full-stack features using Java 8, Spring Framework (Core, MVC, Security), PostgreSQL and Maven, alongside frontend technologies including JavaScript, jQuery, HTML5, Bootstrap, and Stylus.
    \item Built and integrated SOAP-based services to support enterprise system communication and business workflows.
    \item Implemented SAP HCM integrations using SAP JCo and RFC, enabling seamless data exchange between enterprise platforms.
    \item Collaborated in Agile/Scrum teams using Jira and Bitbucket to deliver features iteratively and maintain code quality.
\end{itemize}}

\cventry{Jul 2014 -- Jul 2016}{Research Associate}{Sistema Nacional de Investigadores (CONAHCYT)}{Guanajuato, Mexico}{}{
\begin{itemize}
    \item Appointed as Research Associate within Mexico’s National System of Researchers (CONAHCYT), contributing to academic and applied research initiatives.
    \item Served as professor at the University of Guanajuato (Computer Systems Engineering Faculty), delivering coursework and mentoring undergraduate students.
    \item Recognized as Outstanding Student (2014--2015) for academic excellence and performance.
\end{itemize}}

\cventry{Jul 2013 -- Feb 2014}{Java Developer}{Hermes Transportes Blindados}{Lima, Peru}{}{
\begin{itemize}
    \item Developed enterprise applications and backend components using Java 7, PL/SQL, Spring Core, MyBatis, and Oracle Forms 11g, supporting critical business operations.
    \item Built and integrated SOAP-based services using JAX-WS, enabling reliable communication between enterprise systems.
    \item Deployed and maintained applications on Oracle WebLogic, ensuring stability and performance in production environments.
    \item Designed data models and optimized database interactions using PL/SQL, improving query efficiency and system responsiveness.
    \item Produced technical designs and system documentation using UML, facilitating maintainability and cross-team collaboration.
\end{itemize}}

\cventry{Jul 2012 -- Jun 2013}{Java Developer (Founding Engineer \& Sales Contributor)}{CSTI Corp}{Guayaquil, Ecuador}{}{
\begin{itemize}
    \item Contributed as a founding engineer in early-stage projects, supporting both technical delivery and client-facing activities, including requirements gathering and solution positioning.
    \item Designed and implemented electronic invoicing solutions using Java 6, SOAP, XML, XAdES-BES, XPath, XSD, and XQuery, enabling compliance with regulatory standards.
    \item Built integrations between Java applications and SAP systems using SAP JCo and RFC, facilitating seamless enterprise data exchange.
    \item Developed desktop applications and reporting tools using Java 6 (J2SE, Swing), JDBC, SQLite, iReport, and Apache POI, improving data visualization and operational reporting.
    \item Collaborated with clients and internal teams to deliver tailored solutions, bridging technical implementation and business needs.
\end{itemize}}

\cventry{Sep 2011 -- May 2012}{Java Junior Developer}{GTE Perú}{Lima, Peru}{}{
\begin{itemize}
    \item Developed network management solutions integrating with DSLAM devices (Alcatel, Huawei) using TL-1 and CLI protocols, enabling automated configuration and monitoring of telecommunications infrastructure.
    \item Built web applications and backend components using Java 6 (J2SE), Spring MVC, JSP, MySQL and Servlets, supporting operational workflows and system integrations.
    \item Developed graphical interfaces and visualization tools using Java2D and Ext JS, improving network topology representation and user interaction.
    \item Deployed and maintained applications on Apache Tomcat and Solaris environments, ensuring system stability and performance.
\end{itemize}}

\cventry{Sep 2011 -- May 2012}{Java Junior Developer}{Turismo Dias}{Trujillo, Peru}{}{
\begin{itemize}
    \item Developed web applications using Java 6 (J2SE), JSP, Servlets, and Struts following the MVC pattern, supporting business workflows and system operations.
    \item Built and maintained backend services and data access layers using JPA (EclipseLink) with MySQL and SQL Server, ensuring reliable data persistence and integrity.
    \item Designed and executed SQL queries and database interactions, improving data access efficiency and application performance.
    \item Worked within enterprise development environments using Eclipse, contributing to code quality and maintainability.
\end{itemize}}

\section{Skills}
Java, Kotlin, SQL, JavaScript \\
Microservices Architecture, Spring Boot, Spring Framework, REST APIs, SOAP Services, JPA/Hibernate, Exposed \\
Kafka, RabbitMQ, MQTT, SAP JCo/RFC, API Gateway, Electronic Invoicing Integrations \\
AWS, Docker, Ansible, Terraform, Jenkins, GitLab CI/CD, CI/CD Pipelines \\
PostgreSQL, MySQL, SQL Server, Oracle, H2, SQLite, Apache Ignite \\
Angular, AngularJS, HTML, CSS, JavaScript, JQuery \\
Technical Leadership, Code Reviews, Agile/Scrum, Automated Testing, System Design

\section{Certifications}
AWS Certified Cloud Practitioner \\
Oracle Certified Professional, Java SE 6 Programmer (OCPJP 6) \\
Cisco Certified Network Associate (CCNA 4)

\section{Languages}
Spanish (Native) \\
English (Professional working proficiency)

\section{Education}
Master in Technology Management – University of Guanajuato \\
Computer Science – National University of Trujillo

\end{document}